\documentclass[11pt]{scrartcl}
\usepackage{evan}

\title{Algebraic Combinatorics HW 2}
\author{Evan L, Dallin G, Alexander D, and Sawyer V}
\date{2-13-2024}
\definecolor{palegreen}{rgb}{0.6, 0.98, 0.6}
\begin{document}
\maketitle
\setcounter{section}{1}
\begin{problem}[\textcolor{red}{Random Walks on $\mathbb{Z}$}]
    Consider a random walk on $\mathbb{Z}$ where we start at 0 and move from $i$ to $i + 1$ or $i - 1$ with equal probability.

    \begin{enumerate}[(i)]
        \item Prove that we eventually return to 0 with probability 1.
        \item Compute $a_i$ explicitly and conclude that
        \[
            \sum_{n=0}^{\infty} \frac{a_n}{2_n} = 1
        \]

        What is the sequence ${a_n}$ called in Math literature?
        \item Prove that each number n is visited at least once with probability 1.
        \item Let $H_n$ denote the expected \# steps needed to reach $n$ for the first time. What is wrong with the
        following argument?
        We claim that $H_n = cn$ for some constant $c$. This is true for $n = 0$. So let $n > 0$. On the average, we need $H_1$ steps to reach 1, and then $H_{n-1}$ steps to reach n starting from 1. Hence
        \[
            H_n = 1 + H_{n-1} = c + c(n-1) = cn ; H_1 = c
        \]
    \end{enumerate}
\end{problem}
\begin{proof}
    \begin{enumerate}[(i)]
        \item Let $a_i$ be the number of ways to return to $0$ after $i$ steps.
    \end{enumerate}
\end{proof}
\begin{problem}[\textcolor{red}{Some examples of Hitting times}]
    \begin{enumerate}[(i)]
        \item Find the hitting time between any two vertices of $K_n$.
        \item Find the hitting time between the endpoints of $P_n$ (a path on $n$ vertices).
        \item Find the hitting time between an endpoint of $P_n$ and a vertex at distance $k$ from it.
        \item Find the hitting time between two vertices of $C_n$ (cycle of $n$ vertices) at distance $k$.
        \item Find the hitting time between two `antipodal' vertices of $Q_3$.
    \end{enumerate}
\end{problem}
\begin{problem}
    \begin{enumerate}[(i)]
        \item Show that the following may hold for some graphs $G$ (including regular graphs)
        \[H(u,v)\neq H(v,u),\text{ for some }u,v\in V(G).\]
        \item If $u$ and $v$ have the same degree, then the probability that a random walk starting at $u$ vists $v$ before returning to $u$ is equal to the probability that a random walk starting at $v$ vists $u$ before returning to $u$. What can be said if the degrees of $u$ and $v$ are different?
    \end{enumerate}
\end{problem}
\end{document}
